---
title: "Statistical Learning"
author: "Me"
date: "8/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r}
library(ISLR)
library(dplyr)
library(readr)
library(ggplot2)
library(GGally)
library(mosaic)
library(manipulate)
```


```{r}
data(Auto)

Auto %>%
  slice(1:10)
```



```{r}
dim(Auto)
colnames(Auto)
str(Auto)
```
### Summary statistics
```{r}
summary(Auto)
```

Origin might be better as factor
```{r}
Auto = Auto %>%
  mutate(origin = factor(origin))
summary(Auto)
```

Statistics for a particular variable using favstats
```{r}
Auto %>%
  favstats(~mpg)
```

Or just look at one particular statistic
```{r}
mean(~displacement, data=Auto)
```

## Plotting

```{r}
ggplot(Auto, aes(x = cylinders, y = mpg)) + 
  geom_point()
```
Cylinders are treated as numeric so we convert it to factor

```{r}
Auto = Auto %>%
    mutate(cylinders = factor(cylinders))

ggplot(Auto, aes(x = cylinders, y = mpg)) + 
  geom_point()
```
Produce a boxplot
```{r}
ggplot(Auto) + geom_boxplot(aes(x=cylinders, y=mpg)) + xlab("Cylinders") + ylab("MPG")
```

Create ggpairs
```{r}
Auto %>% select(-cylinders, -origin, -name) %>%
    ggpairs()
```


#College Dataset

```{r}
College = read_csv('College.csv')

College
```


Use summary function to produce numerical summary
```{r}
str(College)

College = College %>%
  mutate(Private = factor(Private))

summary(College)
```

Boxplot of outstate vs private

```{r}
College %>%
  ggplot(aes(Private, Outstate))+
  geom_boxplot()
```

Create a new qualitative variable called Elite by binning the top10perc variable. 

```{r}
College = College %>%
  mutate(Elite = ifelse(College$Top10perc >= 50, 'Yes', 'No'),
         Elite = factor(Elite))

summary(College$Elite)

College %>%
  ggplot(aes(Elite, Outstate))+
  geom_boxplot()
```

### Auto Dataset

Which predictors are quantitative and which are qualtitatve



```{r}
Auto = Auto %>%
  mutate(year = factor(year))
str(Auto)
```

```{r}
Auto_dbl = Auto %>%
  select(mpg, displacement, horsepower, weight, acceleration)

map(Auto_dbl, range)

map(Auto_dbl, mean)

map(Auto_dbl, sd)
```

Now remove observations 10 to 85. What is the range, mean, and sd

```{r}
Auto_subset_dbl = Auto %>%
  slice(-(10:85)) %>%
  select(mpg, displacement, horsepower, weight, acceleration)

map(Auto_subset_dbl, range)

map(Auto_subset_dbl, mean)

map(Auto_subset_dbl, sd)
```

```{r}
ggpairs(Auto %>% select(-name))
```



### Boston Dataset

```{r}
library(MASS)
Boston %>% View()
?Boston
```

Pairwise scatterplots
```{r}
pairs(Boston)
```

```{r}
Boston %>%
  mutate(higher_than_mean = ifelse(crim > mean(crim), 'Yes','No')) %>%
  filter(higher_than_mean=='Yes')
```

How many suburbs are chas == 1

```{r}
Boston %>%
  filter(chas == 1)
```


What is the median pupil teacher ratio among towns

```{r}
median(Boston$ptratio)
```

which suburb has the lowest median owner occupied homes
```{r}
Boston %>%
  filter(age == min(age))
```

















